---
title: "Results graphs and statistical analyses: Experiments 1a, 1b, and 2"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---


# Packages

```{r, message = FALSE, warning=FALSE}
# packages
library(ez)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(overlapping) # to compute the 
library(tidyr)
library(showtext)
#install.packages("expss")
library(expss)
library(dplyr)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```

# Read in data 

Load the data of all studies.

```{r}
# Paths don't have to be adjusted if the data files are in the same folder as the analysis script

# (Exp1a)
tdata_between <- read.delim("data_exp1a.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)

# (Exp1b)
tdata_within <- read.delim("data_exp1b.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)

# (Exp2)
tdata_2d <- read.delim("data_exp2.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)
```


# Create Figures (for Experiment 1a, 1b, and 2)

## Exp. 1a

```{r}
tdata <- tdata_between

# declare factors and their levels 

# Transformation factor
tdata$Transformation <- factor(tdata$Transformation, levels = c("Y_transl", "Small", "180_rot", "90_rot", "X_sym", "Angle", "X_transl"), 
                               labels = c("Y-Tran.", "Size", "180-Rot.", "90-Rot.", "X-Sym.", "Shape", "X-Tran."))


# Counterbalancing factor of "Side"
tdata$Side <- factor(tdata$Side, levels = c("left", "right"), labels = c("left", "right"))


# add selection column coding whether subjects selected the Y-Symmetry or the other transformation
tdata$selection_rec <- c(1:length(tdata$sID))
tdata$selection_rec[tdata$Side =="left" & tdata$Selection == 1] <- "symm"
tdata$selection_rec[tdata$Side =="left" & tdata$Selection == 2] <- "non-symm"
tdata$selection_rec[tdata$Side =="right" & tdata$Selection == 1] <- "non-symm"
tdata$selection_rec[tdata$Side =="right" & tdata$Selection == 2] <- "symm"

# factorize
tdata$selection_rec <- factor(tdata$selection_rec, levels = c("symm", "non-symm"), labels = c("Y-symm.", "other"))
```




```{r}
# create a summary dataset that also contains the percentages (used to create the figure)
plotdata_between <- tdata %>%
  group_by(Transformation, selection_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))
```


```{r, echo = FALSE}
# create a theme to handle the figure design 
myTheme <- theme(plot.title = element_text(face="bold", size = 16),
        axis.title.x = element_text(face = "bold", size = 20),
        axis.title.y = element_text(face = "bold", size = 20),
        axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 18, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        #panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))
```



```{r}

plotdata <- plotdata_between


library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))


ind <- seq(1, nrow(plotdata), by=2)

plotdata_sub <- plotdata[ind, ]


# append CI of the proportions 

plotdata_sub$n <- 122

plotdata_sub$SE <- sqrt((plotdata_sub$pct*(1-plotdata_sub$pct))/plotdata_sub$n)
plotdata_sub$CI <- 1.96 * plotdata_sub$SE


pvalues_x <- c(1:7) # The analyses from which the p-values were obtained come later in the script
pvalues <- c(rep("p < .001",6), "p = .057")

g<- ggplot(plotdata, 
       aes(x = Transformation,
           y = pct,
           fill = selection_rec)) +
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Selected Object",
       x = "Type of Transformation",
       title = "Subjects' Selections in Exp. 1a (N = 854): 3D Context")+
  annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
           ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
           ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
           colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
#ggsave("selections_between.pdf",width=6,height=5)
#ggsave("selections_between.svg",width=6,height=5)
```

## Exp. 1b

```{r}
# Exp. 1b
tdata <- tdata_within

# declare factors and their levels 

# Transformation factor
tdata$Transformation <- factor(tdata$Transformation, levels = c("Y_transl", "Small", "180_rot", "90_rot", "X_sym", "Angle", "X_transl"), 
                               labels = c("Y-Tran.", "Size", "180-Rot.", "90-Rot.", "X-Sym.", "Shape", "X-Tran."))


# Counterbalancing factor of "Side"
tdata$Side <- factor(tdata$Side, levels = c("left", "right"), labels = c("left", "right"))


# add selection column coding whether subjects selected the Y-Symmetry or the other transformation
tdata$selection_rec <- c(1:length(tdata$sID))
tdata$selection_rec[tdata$Side =="left" & tdata$Selection == 1] <- "symm"
tdata$selection_rec[tdata$Side =="left" & tdata$Selection == 2] <- "non-symm"
tdata$selection_rec[tdata$Side =="right" & tdata$Selection == 1] <- "non-symm"
tdata$selection_rec[tdata$Side =="right" & tdata$Selection == 2] <- "symm"

# factorize
tdata$selection_rec <- factor(tdata$selection_rec, levels = c("symm", "non-symm"), labels = c("Y-symm.", "other"))


```


```{r}
# create a summary dataset that also contains the percentages
plotdata_within <- tdata %>%
  group_by(Transformation, selection_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))
```


```{r}
plotdata <- plotdata_within

library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))


ind <- seq(1, nrow(plotdata), by=2)

plotdata_sub <- plotdata[ind, ]


# append CI of the proportions 

plotdata_sub$n <- 122

plotdata_sub$SE <- sqrt((plotdata_sub$pct*(1-plotdata_sub$pct))/plotdata_sub$n)
plotdata_sub$CI <- 1.96 * plotdata_sub$SE



pvalues_x <- c(1:7)
pvalues <- c("p < .001", "p = .001", rep("p < .001",5))


g<- ggplot(plotdata, 
       aes(x = Transformation,
           y = pct,
           fill = selection_rec)) +
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Selected Object",
       x = "Type of Transformation",
       title = "Subjects' Selections in Exp. 1b (N = 854): 3D Context")+
  annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
           ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
           ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
           colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g

#ggsave("selections_within.pdf",width=6,height=5)
#ggsave("selections_within.svg",width=6,height=5)
```



## Exp. 2


```{r}
tdata <- tdata_2d

tdata$Transformation <- factor(tdata$Transformation, levels = c("Y_transl", "Small", "180_rot", "90_rot", "X_sym", "Angle", "X_transl"), 
                               labels = c("Y-Tran.", "Size", "180-Rot.", "90-Rot.", "X-Sym.", "Shape", "X-Tran."))

tdata$Side <- factor(tdata$Side, levels = c("left", "right"), labels = c("left", "right"))


# recode responses so that 1 = first, 2 = second, 3 = both
tdata$selection_rec <- c(1:length(tdata$sID))
tdata$selection_rec[tdata$Side =="left" & tdata$Selection == 1] <- "symm"
tdata$selection_rec[tdata$Side =="left" & tdata$Selection == 2] <- "non-symm"
tdata$selection_rec[tdata$Side =="right" & tdata$Selection == 1] <- "non-symm"
tdata$selection_rec[tdata$Side =="right" & tdata$Selection == 2] <- "symm"



tdata$selection_rec <- factor(tdata$selection_rec, levels = c("symm", "non-symm"), labels = c("Y-symm.", "other"))
```




```{r}
# create a summary dataset that also contains the percentages
plotdata_2d <- tdata %>%
  group_by(Transformation, selection_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))
```




```{r}
plotdata <- plotdata_2d

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

ind <- seq(1, nrow(plotdata), by=2)

plotdata_sub <- plotdata[ind, ]


# append CI of the proportions 

plotdata_sub$n <- 122

plotdata_sub$SE <- sqrt((plotdata_sub$pct*(1-plotdata_sub$pct))/plotdata_sub$n)
plotdata_sub$CI <- 1.96 * plotdata_sub$SE


library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

pvalues_x <- c(1:7)
pvalues <- c("p = .037", "p = .928", "p = .005", "p < .001", "p = .123", "p < .001", "p = .415")


g<- ggplot(plotdata, 
       aes(x = Transformation,
           y = pct,
           fill = selection_rec)) +
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  coord_cartesian(xlim =c(1, 7), ylim = c(0, 1.1))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4.5,
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Selected Object",
       x = "Type of Transformation",
       title = "Subjects' Selections in Exp. 2 (N = 854): 2D Context")+
  annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  annotate("pointrange", x = plotdata_sub$Transformation, y = 1 - plotdata_sub$pct, 
           ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
           ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
           colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  annotate("text", x = pvalues_x, y = Inf, label = pvalues, size = 4, vjust = 1.8)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
        
g
#ggsave("selections_2d.pdf",width=6,height=5)
#ggsave("selections_2d.svg",width=6,height=5)
```


# Create a figure showing the main effect of Dimensions (3D vs. 2D)

```{r}
# combined data frame 

tdata_between$Dimensions <- c(rep("3Da",length(tdata_between$sID)))
tdata_within$Dimensions <- c(rep("3Db",length(tdata_within$sID)))
tdata_2d$Dimensions <- c(rep("2D", length(tdata_2d$sID)))

total <- rbind(tdata_between, tdata_within, tdata_2d)

total$Dimensions <- factor(total$Dimensions, levels = c("3Da", "3Db", "2D"), labels = c("3D (a)", "3D (b)", "2D"))

tdata <- total

tdata$Transformation <- factor(tdata$Transformation, levels = c("Y_transl", "Small", "180_rot", "90_rot", "X_sym", "Angle", "X_transl"), 
                               labels = c("Y-Tran.", "Size", "180-Rot.", "90-Rot.", "X-Sym.", "Shape", "X-Tran."))

tdata$Side <- factor(tdata$Side, levels = c("left", "right"), labels = c("left", "right"))


# recode responses so that 1 = first, 2 = second, 3 = both
tdata$selection_rec <- c(1:length(tdata$sID))
tdata$selection_rec[tdata$Side =="left" & tdata$Selection == 1] <- "symm"
tdata$selection_rec[tdata$Side =="left" & tdata$Selection == 2] <- "non-symm"
tdata$selection_rec[tdata$Side =="right" & tdata$Selection == 1] <- "non-symm"
tdata$selection_rec[tdata$Side =="right" & tdata$Selection == 2] <- "symm"



tdata$selection_rec <- factor(tdata$selection_rec, levels = c("symm", "non-symm"), labels = c("Y-symm.", "other"))

total <- tdata

```


```{r}
# create a summary dataset that also contains the percentages
plotdata_total <- tdata %>%
  group_by(Dimensions, selection_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))
```

```{r, echo = FALSE}
plotdata <- plotdata_total

ind <- seq(1, nrow(plotdata), by=2)

plotdata_sub <- plotdata[ind, ]


# append CI of the proportions 

plotdata_sub$n <- 854

plotdata_sub$SE <- sqrt((plotdata_sub$pct*(1-plotdata_sub$pct))/plotdata_sub$n)
plotdata_sub$CI <- 1.96 * plotdata_sub$SE
```


```{r}
library(scales)
theme_set(theme_light(base_size = 12, base_family = "Poppins"))

g<- ggplot(plotdata, 
       aes(x = Dimensions,
           y = pct,
           fill = selection_rec)) +
  geom_bar(stat = "identity",
           position = "fill") +
  scale_y_continuous(limits = seq(0, 2),
                     breaks = seq(0, 1, .25),
                     expand = c(0,0),
                     label = percent) +
  coord_cartesian(xlim =c(1, 3), ylim = c(0, 1.4))+
  #coord_cartesian(clip = "off")+
  geom_text(aes(label = lbl), 
            size = 4, 
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Percentage", 
       fill = "Selected Object",
       x = "Dimensions",
       title = "Subjects' Selections in 3D vs. 2D Contexts (N = 2562)")+
  #annotate(geom = "hline",yintercept = 0.5, y = 0.5, color = "black", size = 1, linetype='dotted')+
  annotate("pointrange", x = c(1,2,3), y = 1 - plotdata_sub$pct, 
           ymin = 1 - plotdata_sub$pct - plotdata_sub$CI, 
           ymax = 1- plotdata_sub$pct + plotdata_sub$CI, 
           colour = "black", size = 0.8, shape = 22, fill = "lightblue", fatten = 1)+
  annotate("text", x = 1.25, y = Inf, label = "p < .001", size = 4, vjust = 1.6)+
  annotate("text", x = 2.35, y = Inf, label = "p < .001", size = 4, vjust = 4)+
  theme(legend.position = "top", axis.title = element_text(size = 15), axis.text = element_text(size = 13, color = "black"),
        legend.text = element_text(size = 13),legend.title = element_text(size = 13))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g
#ggsave("selections_Dim.pdf",width=6,height=4)
#ggsave("selections_Dim.svg",width=6,height=4)
```

# Statistical tests

## Testing the proportions against chance

### Experiment 1a

```{r}

tdata <- subset(total, Dimensions == "3D (a)")

```


```{r}
# create a summary dataset that also contains the percentages
plotdata_between <- tdata %>%
  group_by(Dimensions, Transformation, selection_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata <- plotdata_between
ind <- seq(1, nrow(plotdata), by=2)

plotdata_sub <- plotdata[ind, ]
```

Binomial-Tests against Chance (50%)

Order of output corresponds to order shown in the figures (Transformations that are plotted on the X-Axis)

```{r}
for (i in plotdata_sub$n){
  print(binom.test(i, 122, p = 0.5, alternative = "two.sided"))
}
```


## Experiment 1b

```{r}

tdata <- subset(total, Dimensions == "3D (b)")

```


```{r}
# create a summary dataset that also contains the percentages
plotdata_within <- tdata %>%
  group_by(Dimensions, Transformation, selection_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata <- plotdata_within
ind <- seq(1, nrow(plotdata), by=2)

plotdata_sub <- plotdata[ind, ]


```
Binomial-Tests against Chance (50%)

Order of output corresponds to order shown in the figures (Transformations that are plotted on the X-Axis)

```{r}
for (i in plotdata_sub$n){
  print(binom.test(i, 122, p = 0.5, alternative = "two.sided"))
}
```
## Experiment 2

```{r}
tdata <- subset(total, Dimensions == "2D")
```


```{r}
# create a summary dataset that also contains the percentages
plotdata_2d <- tdata %>%
  group_by(Dimensions, Transformation, selection_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))



plotdata <- plotdata_2d
ind <- seq(1, nrow(plotdata), by=2)

plotdata_sub <- plotdata[ind, ]
```

Binomial-Tests against Chance (50%)

Order of output corresponds to order shown in the figures (Transformations that are plotted on the X-Axis)

```{r}
for (i in plotdata_sub$n){
  print(binom.test(i, 122, p = 0.5, alternative = "two.sided"))
}

```


# Test the 3D contexts against the 2D context



```{r}
# create a summary dataset that also contains the percentages
plotdata_total <- total %>%
  group_by(Dimensions, selection_rec) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))


```

```{r, echo = FALSE}
plotdata <- plotdata_total

ind <- seq(1, nrow(plotdata), by=2)

plotdata_sub <- plotdata[ind, ]
```



```{r}
test3Da_2D <- prop.test(x = c(plotdata_sub$n[1], plotdata_sub$n[3]), n = c(854, 854))
test3Da_2D
```

```{r}
test3Db_2D <- prop.test(x = c(plotdata_sub$n[2], plotdata_sub$n[3]), n = c(854, 854))
test3Db_2D
```
